EuPPollNet: A European database of plant-pollinator networks

### 2024-08-15 Manuscript/EuPPollNet_Ms.qmd
updates on MS based on coauthors comments

@@

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).

`r format(round(Hymenoptera_percent_metaweb, 2), nsmall = 2)`% of the studies documented interactions from Hymenoptera, `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% from Diptera, `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% from Lepidoptera, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% from Coleoptera.

The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

@@

Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination. EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species. The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids). Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees. Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**). The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`). The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level. Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).

@@

#Coverage
poll_coverage = poll_output$DataInfo$SC *100
plant_coverage = plant_output$DataInfo$SC *100
int_coverage = int_output$DataInfo$SC *100
plant_poll_coverage = plant_poll_output$DataInfo$SC *100

@@

The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_coverage)`% for both taxonomic groups. This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**). The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%. However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**). When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%. The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**). We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**). The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.

@@

Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**). We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**). Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**). Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located in lower latitudes. Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}. Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression. Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network). The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05. Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue). Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database. In addition, the bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}

@@

10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.

>>>>

#Check percentages of hymenoptea groups
hymenoptera_percent_by_group = metadata %>% 
ungroup() %>% 
distinct(Study_id, Taxa_recorded) %>% 
summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))

bee_percent = hymenoptera_percent_by_group$Bee_studies
bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies

#Check percentages of diptera groups
diptera_percent_by_group = metadata %>% 
ungroup() %>% 
distinct(Study_id, Taxa_recorded) %>% 
summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)

syrphid_percent = diptera_percent_by_group$Syrphid_studies
all_diptera_percent = diptera_percent_by_group$Diptera_studies
other_diptera_percent = diptera_percent_by_group$other_diptera_studies

@@

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans. The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

@@

Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination. EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species. The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids). Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees. Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**). The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`). The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level. Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).

@@

#Coverage
poll_sampling_coverage = poll_output$DataInfo$SC *100
plant_sampling_coverage = plant_output$DataInfo$SC *100
int_coverage = int_output$DataInfo$SC *100
plant_poll_coverage = plant_poll_output$DataInfo$SC *100

@@

The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_sampling_coverage)`% for both taxonomic groups. This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**). The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%. However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**). When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%. The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**). We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**). The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.

@@

Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**). We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**). Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**). Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes. Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}. Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression. Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network). The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05. Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue). Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database. The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}

@@

10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located at high elevation areas within Europe.
